---
title: "Optimization-HW2"
author: "Aditya Sathish"
date: "October 23, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##########################################################################################
###Set Up
##########################################################################################
```{r, message = FALSE, warning = FALSE}
rm(list=ls())
require(optrees)
require(igraph)
require(qgraph)
```

##########################################################################################
### Matrix Representation
##########################################################################################

Creating a 1000 by 1000 adjacency matrix
```{r}
n <- 1000
d <- runif(n*n)
d[d < 0.80] <- NA
d <- matrix(d,nrow=n,ncol=n) #reshape the vector

diag(d) <- NA # no self-loops
d[upper.tri(d)] = t(d)[upper.tri(d)]
```


Function that takes in an adjacency matrix and converts it into an adjacency list.
```{r}
AdjMatrix2List <- function (d) {
  ds <- matrix(nrow = (nrow(d)*ncol(d)),ncol = 3, dimnames = list(NULL, c("head", "tail", "weight")))
  c <- 1
  for (i in 1:nrow(d)) {
    for (j in 1:ncol(d)) {
      if (!is.na(d[i,j])) {
        ds[c,] <- c(i, j, d[i,j])
        c <- c + 1
      }
    }
  }
  ds <- ds[1:(c/2),]
  return(ds)
} 
```

Checking above function by converting matrix d to an adjacency list ds and looking at its structure
```{r}
ds <- AdjMatrix2List(d)
str(ds)
```

##########################################################################################
### Euclidean Minimum Spanning Tree Problem
##########################################################################################
Generating 50 random x and y coordinates and plotting them on a graph
```{r}
n <- 50
x <- round(runif(n)*1000)
y <- round(runif(n)*1000)
plot(x,y,pch=16)
```

Creating an adjacency matrix d that calculates the Euclidean distance between every pair of points whose x coordinate is given in x and y coordinate is given in y
```{r}
Points2Matrix <- function (x,y) {
  DAT <- data.frame(x,y, 1:n)
  d <- matrix(nrow = length(x),ncol = length(x))
  for (i in 1:length(x)) {
    for (j in 1:length(y)) {
      d[i,j] <- sqrt(((DAT[i,1] - DAT[j,1])^2) + ((DAT[i,2] - DAT[j,2])^2))
    }
  }
  return(d)
} 

d <- Points2Matrix(x,y)
```
The adjacency matrix (d) created above will be a complete graph as there exists an edge between any two points $d_a$ and $d_b$ in d.  


Converting Adjacency Matrix d to an Adjacency List ds
```{r}
ds <- AdjMatrix2List(d)
```

Running the Miniumum Spanning Tree algorithm using Kruskal

```{r}
ds.mst <- msTreeKruskal(1:n,ds)
```

Plotting line segments that visualize the minimum spanning tree

```{r}
plot.mst <- function (arcList) { 
  for (i in 1:nrow(arcList)) {
    from <- arcList[i,1]
    to <- arcList[i,2]
    xfrom <- x[from]
    xto <- x[to]
    yfrom <- y[from]
    yto <- y[to]
    segments(xfrom, yfrom, xto, yto)
  }
}

plot(x,y, pch = 16)
plot.mst(ds.mst$tree.arcs)
```

##########################################################################################
### Project Scheduling Problem
##########################################################################################
Creating labels and nodes to convert problem to a graph
```{r}
s.labels <- c('a','b','c','d','e','f','g','h','i','j')
s.nodes <- c('90','15','5','20','21','25','14','28','30','45')
```

Manually creating an adjacency list with negative edge weights to convert problem to minimal spanning tree
```{r}
v1 <- c(1,2,-90)
v2 <- c(1,6,-90)
v3 <- c(1,9,-90)
v4 <- c(2,3,-15)
v5 <- c(3,7,-5)
v6 <- c(4,10,-20)
v7 <- c(4,5,-20)
v8 <- c(4,8,-20)
v9 <- c(6,7,-25)
v10 <-c(7,4,-14)
v11 <-c(9,10,-30)

vertices <- rbind(c(v1,v2,v3,v4,v5,v6,v7,v8,v9,v10,v11))
grph <- matrix(vertices, nrow= 11, ncol = 3,byrow = TRUE, dimnames = list(NULL, c("Origin", "Destination", "weight")))
```

Running Belmon-Fords (since edge weights are negative) to calculate shortest possible path
```{r}
bford <- getShortestPathTree(1:length(s.labels),grph,algorithm = "Bellman-Ford", directed = TRUE)
```

Save values of Earliest Start times in a vector called ES. ES takes the negative of bford$distances to convert the numbers back to positive values
```{r}
ES <- -bford$distances
```

Calculating the Earliest Overall Project Completion Time (EOPC)
```{r}
EOPC <- max(ES)+as.numeric(s.nodes[10])
```


Calculating the Earliest Finish Time (EF)
```{r}
EF <- rep(0,length(ES))

for (i in 1:length(ES)){
  EF[i] <- ES[i] + as.numeric(s.nodes[i])
}

```


Creating vertices for transpose of matrix
```{r}
x1 <- c(2,1,-15)
x2 <- c(3,2,-5)
x3 <- c(4,7,-20)
x4 <- c(5,4,-21)
x5 <- c(6,1,-25)
x6 <- c(7,3,-14)
x7<- c(7,6,-14)
x8<- c(8,4,-28)
x9<- c(9,1,-30)
x10<- c(10,9,-45)
x11<- c(10,4,-45)
x12<- c(11,5,0)
x13<- c(11,8,0)
x14<- c(11,10,0)
vertices2 <- rbind(c(x1,x2,x3,x4,x5,x6,x7,x8,x9,x10,x11,x12,x13,x14))
grph2 <- matrix(vertices2, nrow= 14, ncol = 3,byrow = TRUE, dimnames = list(NULL, c("Origin", "Destination", "weight")))
```

Running Belmon-Fords to calculate shortest possible path
```{r}
bford2 <- getShortestPathTree(1:(length(s.labels)+1),grph2,algorithm = "Bellman-Ford", source.node = 11, directed = TRUE)
```

Calculating Latest Finish Times (LF)
```{r}
LF <- bford2$distances + EOPC
LF <- LF[1:10]
```


Calculating Latest Start Time (LS)
```{r}
LS <- rep(0,length(LF))

for (i in 1:length(LF)){
  LS[i] <- LF[i] - as.numeric(s.nodes[i])
}
```


Calculate Slack
```{r}
Slack <- rep(0,length(LF))

for (i in 1:length(LF)){
  Slack[i] <- LF[i] - EF[i]
}
```

Creating a data frame with task names (s.labels), Task Completion Time (s.nodes), Earliest Start Time (ES), Earliest Finish Times (EF), Latest Start Times (LS), Latest Finish Times (LF), and Slack (Slack)
```{r}
df <- data.frame(s.labels, s.nodes, ES, EF, LS, LF, Slack )
```

Converting ES, EF, LS, and LF to dates relative to Nov 1 2017
```{r}
df$ES_dates <- ES + as.Date("2017/11/1")
df$LS_dates <- LS + as.Date("2017/11/1")
df$EF_dates <- EF + as.Date("2017/11/1")
df$LF_dates <- LF + as.Date("2017/11/1")
```

The earliest completion date for the project if it commences on Nov 1, 2017.
```{r}
print(paste("The earliest overall project completion time is", EOPC, "days"))
print(paste("The earliest overall project completion date is", (EOPC + as.Date("2017/11/1"))))
```

The earliest and latest start times for each task which will not alter the completion
date.
```{r}
df[,c(1,8,9)]
```

The earliest and latest finish times for each activity that will not alter the completion
date.

```{r}
df[,c(1,10,11)]
```

The list of tasks which must adhere to a rigid schedule so as not to delay the overall
project, as well as a list of those tasks which have more flexibility.
```{r}
Slack.nodes <- rep(0,sum(Slack == 0))

c <- 1
for (i in 1:length(Slack)){
  if (Slack[i] == 0) {
    Slack.nodes[c] <- s.labels[i]
    c <- c+1
  }
}

print(Slack.nodes)
```



